---
title: "Publication-ready Summary Tables with `{gtsummary}`"
author: "Jubayer Hossain, [Founder & Instructor, CHIRAL Bangladesh](https://chiralbd.org/)"
institute: "Founder, CHIRAL Bangladesh"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    #number_sections: true
    toc_depth: 6
    toc_float: true  
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, warning = FALSE, message = FALSE, fig.align = "center", fig.retina = 5, fig.height = 5, fig.width = 5)
```

# Introduction
The `{gtsummary}` package provides an elegant and flexible way to create publication-ready analytical and summary tables using the R programming language. The `{gtsummary}` package summarizes data sets, regression models, and more, using sensible defaults with highly customizable capabilities.

## `{gtsummary}` Overview
- Create tabular summaries with sensible defaults but highly customizable
- Types of summaries:
  - One-way tables
  - Tow-way tables
  - Cross-tabulation
  - Regression models
  - Survival data
  - Survey data
  - Custom tables
- Report statistics from {gtsummary} tables inline in R Markdown
- Stack and/or merge any table type
- Use themes to standardize across tables
- Choose from different print engines


## Getting Started 
```{r, eval=FALSE}
# Installation
install.packages("gtsummary")

# Install the development version of {gtsummary}
remotes::install_github("ddsjoberg/gtsummary")
```

## Load Packages 
```{r, packages}
# Load packages 
library(tidyverse)
library(gtsummary)
library(kableExtra)
library(gt)
```


```{r}
lungcap <- read.csv("data/LungCapData.csv")
lungcap %>% 
  head() %>% 
  gt() 
```





# Basic Summary Tables using `tbl_summary()`
- Four types of summaries: continuous, continuous2, categorical, and dichotomous
- Statistics are median (IQR) for continuous, n (%) for categorical/dichotomous
- Variables coded 0/1, TRUE/FALSE, Yes/No treated as dichotomous
- Lists NA values under "Unknown"
- Label attributes are printed automatically

## Summary Tables 
```{r}
# Create a basic summary table 
trial %>% 
  select(trt, age, grade, response) %>% 
  tbl_summary()
```


```{r}
# Split table by group 
trial %>% 
  select(trt, age, grade, response) %>% 
  tbl_summary(by = trt)
```

```{r}
# Skip missing data separately 
trial %>% 
  select(trt, age, grade, response) %>% 
  tbl_summary(by = trt, missing = "no")
```

```{r}
# Add column with total number of non-missing observations  
trial %>% 
  select(trt, age, grade, response) %>% 
  tbl_summary(by = trt, missing = "no") %>% 
  add_n()
```


```{r}
# Test for a difference between groups
trial %>% 
  select(trt, age, grade, response) %>% 
  tbl_summary(by = trt, missing = "no") %>% 
  add_n() %>% 
  add_p()
```

```{r}
# Update the column header 
trial %>% 
  select(trt, age, grade, response) %>% 
  tbl_summary(by = trt, missing = "no") %>% 
  add_n() %>% 
  add_p() %>% 
  modify_header(label = "**Variables**")
```

```{r}
# Text formatting: Bold labels 
trial %>% 
  select(trt, age, grade, response) %>% 
  tbl_summary(by = trt, missing = "no") %>% 
  add_n() %>% 
  add_p() %>% 
  modify_header(label = "**Variables**") %>% 
  bold_labels()
```


```{r}
# Text formatting: Italicize Levels
trial %>% 
  select(trt, age, grade, response) %>% 
  tbl_summary(by = trt, missing = "no") %>% 
  add_n() %>% 
  add_p() %>% 
  modify_header(label = "**Variables**") %>% 
  bold_labels() %>% 
  italicize_levels()
```
```{r}
# Table formatting: Pretty print with gt() function
trial %>% 
  select(trt, age, grade, response) %>% 
  tbl_summary(by = trt, missing = "no") %>% 
  add_n() %>% 
  add_p() %>% 
  modify_header(label = "**Variables**") %>% 
  bold_labels() %>% 
  italicize_levels() %>% 
  as_gt() 
```
      
```{r}
# Table formatting: Add title 
trial %>% 
  select(trt, age, grade, response) %>% 
  tbl_summary(by = trt, missing = "no") %>% 
  add_n() %>% 
  add_p() %>% 
  modify_header(label = "**Variables**") %>% 
  bold_labels() %>% 
  italicize_levels() %>% 
  as_gt() %>% 
  gt::tab_header(title = "Table 1. Summary table of 200 patients who received Drug A or Drug B")
```

      
```{r}
# Table formatting: Add Markdown & Bold Title 
trial %>% 
  select(trt, age, grade, response) %>% 
  tbl_summary(by = trt, missing = "no") %>% 
  add_n() %>% 
  add_p() %>% 
  modify_header(label = "**Variables**") %>% 
  bold_labels() %>% 
  italicize_levels() %>% 
  as_gt() %>% 
  gt::tab_header(title = gt::md("**Table 1. Summary table of 200 patients who received Drug A or Drug B**")) 
```


      
```{r}
# Table formatting: Add Markdown & Italic Title 
trial %>% 
  select(trt, age, grade, response) %>% 
  tbl_summary(by = trt, missing = "no") %>% 
  add_n() %>% 
  add_p() %>% 
  modify_header(label = "**Variables**") %>% 
  bold_labels() %>% 
  italicize_levels() %>% 
  as_gt() %>% 
  gt::tab_header(title = gt::md("*Table 1. Summary table of 200 patients who received Drug A or Drug B*")) 
```


```{r}
# Import LungCap Dataset 
pulse_data <- read.csv("data/pulse_data.csv")
pulse_data %>% 
  head() %>% 
  gt() 
```

## Categorical Variable Summary 

```{r}
# Create summary tables
pulse_data %>% 
  select(Gender, Smokes, Alcohol, Exercise, Ran, BMICat) %>% 
  tbl_summary() %>% 
  as_gt()
```

```{r}
# type: specifies the summary type
pulse_data %>% 
  select(Gender, Smokes, Alcohol, Exercise, Ran, BMICat) %>% 
  tbl_summary(type = list(c(Smokes, Alcohol, Exercise, Ran) ~ "categorical")) %>% 
  as_gt()
```


## Numerical Variable Summary Table 
```{r}
pulse_data %>% 
  select(Age, Height, Weight, Pulse1, Pulse2, BMI) %>% 
  tbl_summary(missing = "no", 
              type = everything() ~ "continuous", 
              statistic = all_continuous() ~ "{mean} ± {sd}", 
              digits = all_continuous() ~ 2 
             ) %>% 
  as_gt() 
```

## T-test 

```{r}
pulse_data %>% 
  select(Gender, Age, Height, Weight, Pulse1, Pulse2, BMI) %>% 
  tbl_summary(by = Gender, 
              missing = "no", 
              type = everything() ~ "continuous", 
              statistic = all_continuous() ~ "{mean} ± {sd}", 
              digits = all_continuous() ~ 2 
             ) %>% 
   add_p(
      # perform t-test for all variables
      test = everything() ~ "t.test",
      # assume equal variance in the t-test
      test.args = all_tests("t.test") ~ list(var.equal = TRUE)) %>% 
      modify_header(label = "**Variables**") %>% 
      bold_labels() %>%
      italicize_levels() %>% 
      as_gt() %>% 
      gt::tab_header(
      title = gt::md("**T-test Table**")
)
  
```


## ANOVA 

```{r}
pulse_data %>% 
  select(BMICat, Age, Height, Weight, Pulse1, Pulse2, BMI) %>% 
  tbl_summary(by = BMICat, 
              missing = "no", 
              type = everything() ~ "continuous", 
              statistic = all_continuous() ~ "{mean} ± {sd}", 
              digits = all_continuous() ~ 2 
             ) %>% 
   # Perform One-way ANOVA 
   add_p(test = everything() ~ "aov") %>% 
      modify_header(label = "**Variables**") %>% 
      bold_labels() %>%
      italicize_levels() %>% 
      as_gt() %>% 
  gt::tab_header(
      title = gt::md("**ANOVA Table**")
)
```


## Univariate Regression

```{r}
trial %>%
  select(response, age, grade) %>%
  tbl_uvregression(
    method = glm,
    y = response,
    method.args = list(family = binomial),
    exponentiate = TRUE,
    pvalue_fun = ~style_pvalue(.x, digits = 2)
  ) %>%
  add_global_p() %>%  # add global p-value 
  add_nevent() %>%    # add number of events of the outcome
  add_q() %>%         # adjusts global p-values for multiple testing
  bold_p() %>%        # bold p-values under a given threshold (default 0.05)
  bold_p(t = 0.10, q = TRUE) %>% # now bold q-values under the threshold of 0.10
  bold_labels()
```




