---
title: "Data Visualization with ggplot2"
author: "Jubayer Hossain, [Founder & Instructor, CHIRAL Bangladesh](https://chiralbd.org/)"
institute: "Founder, CHIRAL Bangladesh"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    #number_sections: true
    toc_depth: 6
    toc_float: true 
    theme: flatly
--- 

# What is ggplot2?
- ggplot2 is a data visualization package for the statistical programming language R.
- Created by Hadley Wickham in 2005, ggplot2 is an implementation of Leland Wilkinson’s Grammar of Graphics—a general scheme for data visualization which breaks up graphs into semantic components such as scales and layers.


```{r pkgs}
library(tidyverse)
```


# Dataset and Data Exploration 
```{r data}
pulse_data <- read.csv("data/pulse_data.csv")
```

# Distribution Plots 

## Violin
An alternative to boxplot to compare the distribution of several groups.

- Violinplots allow to visualize the distribution of a numeric variable for one or several groups.

- It is really close from a boxplot, but allows a deeper understanding of the distribution.

- Violins are particularly adapted when the amount of data is huge and showing individual observations gets impossible. 

```{r violin-plot}
ggplot(pulse_data, aes(x=BMICat, y=Age, fill=Gender)) + # fill=name allow to automatically dedicate a color for each group
  geom_violin()
```


## Density

## Histogram

## Boxplot

# Correlation Plots 







