# Reshaping Data

```{r, include=FALSE}
# packages 
library(tidyverse)
library(readxl)
```

## Reshaping Data - Change the Layout of a Dataset 

Reshaping data involves changing the layout of a dataset to make it more suitable for analysis or presentation. This typically includes transforming data between wide and long formats or converting between different types of data structures.

## Reshaping from Wide to Long Format
- `tidyr::gather()`: Converts wide data into long format by gathering columns into key-value pairs.
- `pivot_longer()`: Similar to `gather()`, it transforms wide data into long format but provides more flexibility and options.

```{r, wide-long}
wide_data <- read_excel("data/Life_Expectancy_Wide.xlsx")

# convert into long 
long_data <- wide_data %>% 
  tidyr::pivot_longer(cols = 2:75, names_to = "Year", values_to = "LifeExp")
head(long_data)
```

## Reshaping from Long to Wide Format
- `tidyr::spread()`: Converts long data into wide format by spreading values from a column into multiple columns based on unique identifiers.
- `pivot_wider()`: Similar to spread(), it transforms long data into wide format but provides more flexibility and options.

```{r, long-wide}
# convert into wide 
wide_data <- long_data %>% 
  pivot_wider(names_from = "Year", values_from = "LifeExp")
head(wide_data)
```